---
title: "COPD exacerbation"
author: "George Qian"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

31678
Prospective cohort study of adults hospitalised
with acute COPD exacerbations: SARS-CoV-2-
associated patient characteristics and clinical
outcomes


Background
Acute exacerbations of chronic obstructive pulmonary disease (AECOPD) are common and associated with
respiratory function decline, increased morbidity and mortality. It is unclear how COVID19 has impacted
AECOPD phenotype and if it is associated with worse outcomes compared to other infections in COPD
patients.

Aims
To explore changes in AECOPD clinical characteristics and outcomes during the COVID-19 pandemic.

Methods
A prospective cohort study of all adults ≥18y admitted to either acute care hospital in Bristol UK, Aug 20–Jul
21. Patients presenting with signs/symptoms or a clinical/radiological diagnosis of acute respiratory disease
were included.

Results
2557 hospitalisations with AECOPD were recorded (incidence 361 per 100000); 13% had SARS-CoV-2
identified, 69% had another pathogen and 18% had no infectious aetiology identified. Patient characteristics
and clinical features were similar for patients with and without COVID19.

ICU admissions were more common with SARS-CoV-2 than other infections (4% v 1%, P<0.001). SARS-
CoV-2 associated AECOPD had greater inpatient (25% v 9%, P<0.001) and 30-day mortality (25% v 12%,
P<0.001) than AECOPD associated with other infections. Among all AECOPD patients who had received
≥1 dose of any COVID vaccination, ICU admission rates were lower than for unvaccinated persons (4% v
0%, P<0.01).

Discussion
SARS-CoV-2 infection was associated with worse outcomes among persons hospitalized with AECOPD.
SARS-CoV-2 vaccination was effective in reducing disease severity and ICU admission. More data are
needed about the effectiveness of COVID19 vaccines for modifying disease severity in COPD patients.

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = TRUE,
  warning = TRUE
)

# The `here` library sorts out all relative paths.
here::i_am("copd-exacerbation/copd-exacerbation.Rmd")

# common-setup will attempt to load all the dependencies including 
# ggrrr and dtrackr.
source(here::here("common-setup.R"))

# This creates a function that can be used to put all the outputs (like) a specific directory
out = outputter(directory=here::here("output/copd-exacerbation"),datedSubdirectory = TRUE)

# don't need this yet.
# sup = data_supplement(out = out)
# options(hide.supplementary.tables=TRUE)

# this will set the latest date of a data file that will be loaded by the data load scripts.
# this forces a particular version fo the files for this script.
options(reproduce.at="2022-06-21")

# ignore these options for the time being they are
# for formatting the flowchart.
options(dtrackr.strata_glue="{.value}")
options(dtrackr.strata_sep=", ")
options(dtrackr.default_message="{.count} admissions")
options(dtrackr.show_zero_exclusions=FALSE)
options(dtrackr.default_count_subgroup = "{.name}: {.count} admissions")

```

## Load data

```{r data}

# Load the data from input directory. 
# This assumes a file will exist that starts with AvonCAPLRTDCentralDa_y1..., AvonCAPLRTDCentralDa_y2... in the input directory. _y1, _y2 must be present.
# it will produce one data set from the multiple files.
lrtd_data = load_data("AvonCAPLRTDCentralDa")

# This will rename the variables and sort out factor levels. 
# This has to be done for the ltrd data in a slightly bespoke way because of the multiple file formats.
lrtd_norm = normalise_data(lrtd_data)
# This will augment the data with "standard" outcomes that I have defined in the data-wrangling.R file.
# This augments the data with a set of "standard" extra variables. However some of these require
# precise dates that are not present in the lrtd data sets and so this throws a set of warnings around
# not being able to compute things. I need to tidy this up, and am thinking about a naming convention for these things that
# are data set dependent.
lrtd_aug = lrtd_norm %>% augment_data()


# this constructs a helper list which tells you the dataframe columns and factor levels as a list
v = lrtd_aug %>% get_value_sets()
# typing `v$` at the console should give you some auto-complete support to help navigate the copd_study dataframe.

```

## Study specific pipeline

```{r}
# Might want to change these which are defaulted to the start and end of the data:
start_week = min(lrtd_aug$admission.study_week,na.rm=TRUE)
end_week = max(lrtd_aug$admission.study_week,na.rm=TRUE)
reproduce_at = as.Date(getOption("reproduce.at"))
# maybe for example:
# start_week = study_week(as.Date("2021-01-01"))
# end_week = study_week(as.Date("2022-01-01"))

# This sets up the data pipeline
# The data frame is being tracked using dtrackr (https://terminological.github.io/dtrackr/index.html)
# It also does some standard exclusions. 
# Depending on what you want to do these exclusions may need to be reviewed.
copd_study = lrtd_aug %>% 
  mutate(
    admission.copd_exacerbation_status = case_when(
      diagnosis.SOC_exacerbation_COPD == v$diagnosis.SOC_exacerbation_COPD$yes ~ "COPD Exacerbation",
      TRUE ~ "Not COPD Exacerbation"
    ) %>% factor(levels=c("Not COPD Exacerbation","COPD Exacerbation"))
  ) %>%
  include_any(
    admission.study_week >= start_week & admission.study_week < end_week ~ "{.count} patients admitted \nbetween {start_date_of_week(start_week)} \n and {start_date_of_week(end_week)}"
  ) %>%
  exclude_all(
      admin.consented == v$admin.consented$`Declined consent` | admin.pp_consented == v$admin.pp_consented$`Declined consent` ~ "{.excluded} declined consent",
      admin.withdrawal == v$admin.withdrawal$yes ~ "{.excluded} withdrew consent",
      .stage = "consent exclusions"
  ) %>%
  comment(.messages = "{.count} patients") %>%
  exclude_all(
      demog.age<18 ~ "{.excluded} under 18 on admission",
      # Taking this out as quite a lot. maybe COPD pts are always symptomatic
      # admission.duration_symptoms > 10 ~ "{.excluded} symptomatic >10 days before admission",
      admission.date > reproduce_at-7 ~ "{.excluded} with admission after {fdmy(reproduce_at-7)}",
      admission.episode > 1 ~ "{.excluded} repeat admissions",
      is.na(comorbid.copd) ~ "{.excluded} with unknown COPD status",
      admission.copd_exacerbation_status == "COPD Exacerbation" & comorbid.copd == v$comorbid.copd$no ~ "{.excluded} COPD exacerbations with negative COPD status",
      admission.episode > 1 ~ "{.excluded} repeat admissions",
      admin.duplicate == v$admin.duplicate$yes ~ "{.excluded} records with identical record number and admission date",
      .stage = "standard exclusions"
  ) %>%
  group_by(admission.copd_exacerbation_status, .messages="with comorbid COPD") %>%
  
  comment("{.count} patients") %>%
  dtrackr::count_subgroup(.subgroup = genomic.variant_inferred, .headline = "variant") %>%
  dtrackr::count_subgroup(.subgroup = admission.category, .headline = "cause")
# N.B. more data filtering usually expected here. 
# probably want to include_any with COPD 


# This will give a xlsx of all the cases that were excluded
# copd_study %>% excluded() %>% openxlsx::write.xlsx(file = out("exclusions.xlsx"))

# This will give you a figure 1 consort flow chart. This is a dtrackr function.
copd_study %>% flowchart(filename = out("fig1-flowchart"))

# Untracking the data frame means that normal dplyr / tidyverse functions will be used rather than 
# the overloaded versions from dtrackr. This will be slightly more reliable, but from this point onwards no changes will be tracked.
copd_data = copd_study %>% untrack()
copd_only = copd_study %>% filter(comorbid.copd == v$comorbid.copd$yes)
```

## Plotting stuff

```{r}

# ggrrr has a set of functions making ggplots behave themselves.
# a prejudiced theme for publication and sensible size defaults are enforced by the following
ggrrr::gg_pedantic()

p = ggplot(copd_study,aes(x=comorbid.copd,fill=comorbid.copd))+geom_bar()

# ggrrr has some ggplot saving functions that enforces standard sizes and saves in multiple formats:
# this is good for knowing what it is going to look like at publication. Overriding R markdowns decision making about sizing.
p %>% ggrrr::gg_save_as(out("fig2-random-plot"),size = ggrrr::std_size$sixth)
```

## Comparison tables

```{r}

# I've looked at a set of functions to generate comparison tables
# there are in descriptive-tables.R

t = compare_population(copd_only, 
    populationVar = admission.category,
    comparisonVars = vars(
      demog.age,
      demog.gender,
      demog.ethnicity,
      admission.cci_category,
      demog.imd_decile,
      demog.smoker,
      admission.on_inhaled_corticosteroids
    ),
    p.value=TRUE,na.rm = TRUE
) %>% population_comparison_table(compact = TRUE,show_method = FALSE)

# col=1
# rows = (1:nrow(t))[!t %>% huxtable::header_rows()]
# spans = attributes(t)$rowspan[rows,col]
# toadj = (rows+spans)[rows+spans > cummax(lag(rows+spans,default = 0))]
# toadj = rev(as.integer(names(toadj)))
# t2 = t
# for (row in toadj) {
#   t2 = huxtable::insert_row(ht = t2, t[row,1:col],fill = t[row,col], after=row-1)
#   t2 = t2 %>% set_colspan(row = row, col = col, value = ncol(t)-col+1)
#   t2 = t2 %>% set_bottom_border(row = row, col = col:ncol(t), value = 0)
#   t2[row+1,col]=""
# }
# headers = (1:nrow(t))[t %>% huxtable::header_rows()]
# t2[headers,col] = ""
#   

t2 = t %>% hux_nest_group(col=1)

frac = function(x) return(x/sum(x,na.rm = TRUE))
# ggrrr has some functions to render tables into different formats. 
# although docx seems not to be working 
t2 %>% hux_set_font(defaultFontSize = 6) %>% huxtable::set_col_width(frac(c(0.2,0.8,1,1,1,0.4))) %>% hux_save_as(out("tab1-random-comparison"),formats = c("html","pdf","png"))

```
