---
title: "COPD exacerbation"
author: "George Qian"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

31678
Prospective cohort study of adults hospitalised
with acute COPD exacerbations: SARS-CoV-2-
associated patient characteristics and clinical
outcomes


Acute COPD exacerbation is a key and independent risk factor for new
cardiovascular events within 30 days in a hospitalised cohort.

Are patients with an exacerbation of COPD at a higher risk of cardiovascular
events than those presenting with a LRTI without a history of COPD, in a
hospitalised cohort? (Description as a high-risk group- shared risk factors with
CVD: hypercholesterolaemia, hypertension, Diabetes, peripheral arterial disease,
smoking) Which risk factors are specific to COPD? Outcome of interest:
Cardiovascular event: New AF, STEMI, NSTEMI, arrythmia, HF (not on admission,
but defined by subsequent diagnosis in hospital or by GP), PE/DVT, Stroke

Disease (exposure): COPD – physician diagnosed (exacerbation severity defined by
DECAF score but no means to define severity of baseline disease)

# Secondary analysis:

Does the nature of the COPD exacerbation (infective vs non-infective) influence
cardiovascular outcomes? (Infective exacerbation COPD – pneumonic changes on
imaging, CRP>100, CURB-65, COVID confirmed) Are serum biomarkers (admission
troponin, admission BNP, CRP) a useful tool in predicting cardiovascular events
in the COPD cohort?
 




```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = TRUE,
  warning = TRUE
)

# The `here` library sorts out all relative paths.
here::i_am("copd-exacerbation/copd-exacerbation.Rmd")

# common-setup will attempt to load all the dependencies including 
# ggrrr and dtrackr.
source(here::here("common-setup.R"))

# This creates a function that can be used to put all the outputs (like) a specific directory
out = outputter(directory=here::here("output/copd-exacerbation"),datedSubdirectory = TRUE)

# don't need this yet.
# sup = data_supplement(out = out)
# options(hide.supplementary.tables=TRUE)

# this will set the latest date of a data file that will be loaded by the data load scripts.
# this forces a particular version fo the files for this script.
options(reproduce.at=NULL)

# ignore these options for the time being they are
# for formatting the flowchart.
options(dtrackr.strata_glue="{.value}")
options(dtrackr.strata_sep=", ")
options(dtrackr.default_message="{.count} admissions")
options(dtrackr.show_zero_exclusions=FALSE)
options(dtrackr.default_count_subgroup = "{.name}: {.count} admissions")

```

## Load data

```{r data}

# meta = load_data("metadata")
# meta2 = meta %>% mutate(label1 = stringr::str_remove_all(`Field Label`, "<[^>]+>") %>% trimws())
# meta2 = meta2 %>% mutate(levels = `Choices, Calculations, OR Slider Labels` %>% 
#                            stringr::str_split("\\s?\\|\\s?") %>% 
#                            purrr::map(~ tibble(level = .x %>% stringr::str_remove("^[^,]+,\\s?"), label=.x %>% stringr::str_remove(",\\s?.*$"))))
# meta3 = meta2 %>% unnest(levels) %>% mutate(level = stringr::str_remove_all(level, "<[^>]+>") %>% trimws())

```


```{r}
# Load the data from input directory. 
# This assumes a file will exist that starts with AvonCAPLRTDCentralDa_y1..., AvonCAPLRTDCentralDa_y2... in the input directory. _y1, _y2 must be present.
# it will produce one data set from the multiple files.
lrtd_data = load_data("AvonCAPLRTDCentralDa",merge = TRUE)

# This will rename the variables and sort out factor levels. 
# This has to be done for the lrtd data in a slightly bespoke way because of the multiple file formats.

lrtd_norm = normalise_data(lrtd_data)
# This will augment the data with "standard" outcomes that I have defined in the data-wrangling.R file.
# This augments the data with a set of "standard" extra variables. However some of these require
# precise dates that are not present in the lrtd data sets and so this throws a set of warnings around
# not being able to compute things. I need to tidy this up, and am thinking about a naming convention for these things that
# are data set dependent.
lrtd_aug = lrtd_norm %>% augment_data()


# this constructs a helper list which tells you the dataframe columns and factor levels as a list
v = lrtd_aug %>% get_value_sets()
# typing `v$` at the console should give you some auto-complete support to help navigate the copd_study dataframe.

```

## Study specific pipeline

```{r}
# Might want to change these which are defaulted to the start and end of the data:
start_week = min(lrtd_aug$admission.study_week,na.rm=TRUE)
end_week = max(lrtd_aug$admission.study_week,na.rm=TRUE)
reproduce_at = as.Date(getOption("reproduce.at",default = Sys.Date()))
# maybe for example:
# start_week = study_week(as.Date("2021-01-01"))
# end_week = study_week(as.Date("2022-01-01"))

# This sets up the data pipeline
# The data frame is being tracked using dtrackr (https://terminological.github.io/dtrackr/index.html)
# It also does some standard exclusions. 
# Depending on what you want to do these exclusions may need to be reviewed.
copd_study = lrtd_aug %>% 
  mutate(
    admission.copd_exacerbation_status = case_when(
      diagnosis.SOC_exacerbation_COPD == v$diagnosis.SOC_exacerbation_COPD$yes ~ "COPD Exacerbation",
      TRUE ~ "Not COPD Exacerbation"
    ) %>% factor(levels=c("Not COPD Exacerbation","COPD Exacerbation"))
  ) %>%
  include_any(
    admission.study_week >= start_week & admission.study_week < end_week ~ "{.count} patients admitted \nbetween {start_date_of_week(start_week)} \n and {start_date_of_week(end_week)}"
  ) %>%
  exclude_all(
      admin.consented == v$admin.consented$`Declined consent` | admin.pp_consented == v$admin.pp_consented$`Declined consent` ~ "{.excluded} declined consent",
      admin.withdrawal == v$admin.withdrawal$yes ~ "{.excluded} withdrew consent",
      .stage = "consent exclusions"
  ) %>%
  comment(.messages = "{.count} patients") %>%
  exclude_all(
      demog.age<18 ~ "{.excluded} under 18 on admission",
      # Taking this out as quite a lot. maybe COPD pts are always symptomatic
      # admission.duration_symptoms > 10 ~ "{.excluded} symptomatic >10 days before admission",
      admission.date > reproduce_at-7 ~ "{.excluded} with admission after {.fdmy(reproduce_at-7)}",
      admission.episode > 1 ~ "{.excluded} repeat admissions",
      is.na(comorbid.copd) ~ "{.excluded} with unknown COPD status",
      admission.copd_exacerbation_status == "COPD Exacerbation" & comorbid.copd == v$comorbid.copd$no ~ "{.excluded} COPD exacerbations with negative COPD status",
      admission.episode > 1 ~ "{.excluded} repeat admissions",
      admin.duplicate == v$admin.duplicate$yes ~ "{.excluded} records with identical record number and admission date",
      .stage = "standard exclusions"
  ) %>%
  group_by(admission.copd_exacerbation_status, .messages="with comorbid COPD") %>%
  
  comment("{.count} patients") %>%
  dtrackr::count_subgroup(.subgroup = genomic.variant_inferred, .headline = "variant") %>%
  dtrackr::count_subgroup(.subgroup = admission.category, .headline = "cause")
# N.B. more data filtering usually expected here. 
# probably want to include_any with COPD 


# This will give a xlsx of all the cases that were excluded
# copd_study %>% excluded() %>% openxlsx::write.xlsx(file = out("exclusions.xlsx"))

# This will give you a figure 1 consort flow chart. This is a dtrackr function.


# Untracking the data frame means that normal dplyr / tidyverse functions will be used rather than 
# the overloaded versions from dtrackr. This will be slightly more reliable, but from this point onwards no changes will be tracked.
copd_data = copd_study %>% untrack()
copd_only = copd_study %>% filter(comorbid.copd == v$comorbid.copd$yes)
```

## Plotting stuff

```{r}

# ggrrr has a set of functions making ggplots behave themselves.
# a prejudiced theme for publication and sensible size defaults are enforced by the following
ggrrr::gg_pedantic()

v = copd_study %>% ggrrr::get_value_sets()
copd_study2 = copd_study %>% mutate(
    admission.copd_ilrtd_status = case_when(
      admission.copd_exacerbation_status == v$admission.copd_exacerbation_status$`COPD Exacerbation` & admission.category == v$admission.category$`Confirmed SARS-CoV-2` ~ "COVID with COPD",
      admission.copd_exacerbation_status == v$admission.copd_exacerbation_status$`COPD Exacerbation` & admission.category == v$admission.category$`No evidence SARS-CoV-2` ~ "Other iLRTD with COPD",
      admission.copd_exacerbation_status == v$admission.copd_exacerbation_status$`Not COPD Exacerbation` & admission.category == v$admission.category$`Confirmed SARS-CoV-2` ~ "COVID without COPD",
      admission.copd_exacerbation_status == v$admission.copd_exacerbation_status$`Not COPD Exacerbation` & admission.category == v$admission.category$`No evidence SARS-CoV-2` ~ "Other iLRTD without COPD",
      TRUE ~ NA_character_
    ) %>% factor(levels = c("COVID with COPD","Other iLRTD with COPD","COVID without COPD","Other iLRTD without COPD")),
    admission.ilrtd_status = case_when(
      admission.category == v$admission.category$`Confirmed SARS-CoV-2` ~ "COVID",
      admission.category == v$admission.category$`No evidence SARS-CoV-2` ~ "Other iLRTD",
      TRUE ~ NA_character_
    ) %>% factor(levels = c("COVID","Other iLRTD"))
  ) %>% 
  exclude_all(is.na(admission.ilrtd_status) ~ "{.excluded} non infective causes of aLRTD") %>%
  dtrackr::comment("{.count} patients")

copd_study2 %>% flowchart(filename = out("fig1-flowchart"))

# p = ggplot(copd_study,aes(x=comorbid.copd,fill=comorbid.copd))+geom_bar()

# ggrrr has some ggplot saving functions that enforces standard sizes and saves in multiple formats:
# this is good for knowing what it is going to look like at publication. Overriding R markdowns decision making about sizing.
# p %>% ggrrr::gg_save_as(out("fig2-random-plot"),size = ggrrr::std_size$sixth)
```

## Comparison tables

```{r}

# I've looked at a set of functions to generate comparison tables
# there are in descriptive-tables.R

t1 = compare_population(copd_study2, 
    populationVar = admission.copd_exacerbation_status,
    comparisonVars = vars(
      demog.age,
      demog.gender,
      demog.ethnicity,
      admission.cci_category,
      demog.imd_decile,
      demog.smoker,
      admission.on_inhaled_corticosteroids
    ),
    p.value=TRUE,na.rm = TRUE
) 

t1 = t1 %>% population_comparison_table(compact = TRUE,show_method = FALSE)

covid_positive_subgroup = copd_study2 %>% filter(admission.ilrtd_status == "COVID")

t2 = compare_population(covid_positive_subgroup, 
    populationVar = admission.copd_ilrtd_status,
    comparisonVars = vars(
      demog.age,
      demog.gender,
      demog.ethnicity,
      admission.cci_category,
      demog.imd_decile,
      demog.smoker,
      admission.on_inhaled_corticosteroids
    ),
    p.value=TRUE,na.rm = TRUE
) 

t2 = t2 %>% population_comparison_table(compact = TRUE,show_method = FALSE)

covid_negative_subgroup = copd_study2 %>% filter(admission.ilrtd_status == "Other iLRTD")


t3 = compare_population(covid_negative_subgroup, 
    populationVar = admission.copd_ilrtd_status,
    comparisonVars = vars(
      demog.age,
      demog.gender,
      demog.ethnicity,
      admission.cci_category,
      demog.imd_decile,
      demog.smoker,
      admission.on_inhaled_corticosteroids
    ),
    p.value=TRUE,na.rm = TRUE
) 

t3 = t3 %>% population_comparison_table(compact = TRUE,show_method = FALSE)

t = t1 %>% huxtable::add_columns(t2[3:5]) %>% huxtable::add_columns(t3[3:5]) %>% hux_nest_group(col=1)

t %>% hux_set_font(defaultFontSize = 6) %>%
  hux_save_as(out("tab1-demographic-comparison"),formats = c("html","pdf","png"))
```


```{r}

# I've looked at a set of functions to generate comparison tables
# there are in descriptive-tables.R

t1 = compare_population(copd_study2, 
    populationVar = admission.copd_exacerbation_status,
    comparisonVars = vars(
      outcome.STEMI,
      outcome.NSTEMI,
      outcome.PE,
      outcome.DVT,
      outcome.stroke,
      outcome.new_AF,
      outcome.heart_failure,
      outcome.acute_renal_failure,
      outcome.no_complications
    ),
    p.value=TRUE,na.rm = TRUE
) 

t1 = t1 %>% population_comparison_table(compact = TRUE,show_method = FALSE)


t2 = compare_population(covid_positive_subgroup, 
    populationVar = admission.copd_ilrtd_status,
    comparisonVars = vars(
      outcome.STEMI,
      outcome.NSTEMI,
      outcome.PE,
      outcome.DVT,
      outcome.stroke,
      outcome.new_AF,
      outcome.heart_failure,
      outcome.acute_renal_failure,
      outcome.no_complications
    ),
    p.value=TRUE,na.rm = TRUE
) 

t2 = t2 %>% population_comparison_table(compact = TRUE,show_method = FALSE)


t3 = compare_population(covid_negative_subgroup, 
    populationVar = admission.copd_ilrtd_status,
    comparisonVars = vars(
      outcome.STEMI,
      outcome.NSTEMI,
      outcome.PE,
      outcome.DVT,
      outcome.stroke,
      outcome.new_AF,
      outcome.heart_failure,
      outcome.acute_renal_failure,
      outcome.no_complications
    ),
    p.value=TRUE,na.rm = TRUE
) 

t3 = t3 %>% population_comparison_table(compact = TRUE,show_method = FALSE)

t = t1 %>% huxtable::add_columns(t2[3:5]) %>% huxtable::add_columns(t3[3:5]) %>% hux_nest_group(col=1)

t %>% hux_set_font(defaultFontSize = 6) %>%
  hux_save_as(out("tab2-outcome-comparison"),formats = c("html","pdf","png"))



```


```{r}

copd_study2 = copd_study2 %>% mutate(across(.cols = where(is.ordered), .fns = factor, ordered=FALSE))

allVars = vars(
      admission.ilrtd_status,    
      admission.copd_exacerbation_status,
      demog.age,
      demog.gender,
      demog.ethnicity,
      admission.cci_category,
      demog.imd_decile,
      demog.smoker,
      admission.on_inhaled_corticosteroids)

copd_impute = avoncap:::impute_data(copd_study2, allVars)

```


```{r}
form = . ~ admission.ilrtd_status + admission.copd_exacerbation_status+demog.age+
      demog.gender+demog.ethnicity+admission.cci_category+demog.imd_decile+
      demog.smoker+admission.on_inhaled_corticosteroids

modelConfig = tibble::tribble(
    ~modelName, ~form,
    "STEMI", update(form,outcome.STEMI ~ .),
    "NSTEMI", update(form,outcome.NSTEMI ~ .),
    "PE", update(form,outcome.PE ~ .),
    "DVT", update(form,outcome.DVT ~ .),
    "Stroke", update(form,outcome.stroke ~ .),
    "New AF", update(form,outcome.new_AF ~ .),
    "HF", update(form,outcome.heart_failure ~ .),
    "ARF", update(form,outcome.acute_renal_failure ~ .)
)

copd_qp_model = avoncap:::run_models(copd_impute, modelConfig, avoncap:::logistic_regression)

combinedTable = copd_qp_model %>% summarise_boots(allVars, statistic = "OR") 

combinedTable  %>% hux_auto_widths() %>%
  hux_save_as(out("tab3-relative-risk-models"),formats = c("html","pdf","png"))

# tmp = mice::complete(copd_impute,1)
# tmp %>% group_by(admission.ilrtd_status) %>% count()

```

```{r}
form = . ~ admission.copd_ilrtd_status + demog.age+
      demog.gender+demog.ethnicity+admission.cci_category+demog.imd_decile+
      demog.smoker+admission.on_inhaled_corticosteroids

modelConfig = tibble::tribble(
    ~modelName, ~form,
    "STEMI", update(form,outcome.STEMI ~ .),
    "NSTEMI", update(form,outcome.NSTEMI ~ .),
    "PE", update(form,outcome.PE ~ .),
    "DVT", update(form,outcome.DVT ~ .),
    "Stroke", update(form,outcome.stroke ~ .),
    "New AF", update(form,outcome.new_AF ~ .),
    "HF", update(form,outcome.heart_failure ~ .),
    "ARF", update(form,outcome.acute_renal_failure ~ .)
)

copd_qp_model = avoncap:::run_models(copd_impute, modelConfig, avoncap:::logistic_regression)

combinedTable = copd_qp_model %>% summarise_boots(vars(
      admission.copd_ilrtd_status,    
      demog.age,
      demog.gender,
      demog.ethnicity,
      admission.cci_category,
      demog.imd_decile,
      demog.smoker,
      admission.on_inhaled_corticosteroids), statistic = "OR") 

combinedTable  %>% hux_auto_widths() %>%
  hux_save_as(out("tab4-relative-risk-interaction-models"),formats = c("html","pdf","png"))

# tmp = mice::complete(copd_impute,1)
# tmp %>% group_by(admission.ilrtd_status) %>% count()

```