---
title: "Vaccine induce and naturally acquired antibody to SARS-CoV-2"
author: "Rob Challen"
date: "24/06/2022"
output: html_document
knit: ggrrr::knit_versioned("html","output/omicron-severe-disease-suppl")
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = TRUE,
  warning = TRUE
)

# The `here` library sorts out all relative paths.
here::i_am("vaccine-effectiveness/waning-immunity.Rmd")

# common-setup will attempt to load all the dependencies including 
# ggrrr and dtrackr.
source(here::here("common-setup.R"))
source(here::here("estimators.R"))

# This creates a function that can be used to put all the outputs (like) a specific directory
out = outputter(directory=here::here("output/vaccine-effectiveness"))

```



```{r cars}
# serology = readr::read_csv(unz(here::here("input/week15_2022_seroprev_modellers.csv.zip")))
serology = readr::read_csv(here::here("input/week15_2022_seroprev_modellers.csv"))
serology = serology %>% track()
serology = serology %>% mutate(across(ends_with("date"), .fns = ~ as.Date(.x, "%d%b%Y")))

```
# N-antibody community level

* Log link age and date linear model.
* Naturally acquired immunity

```{r}
model = locfit::locfit(RocheN_units ~locfit::lp(age,SampleDate,nn=0.2,scale=TRUE),data = serology, link="log")
summary(model)
new = tidyr::crossing(age = 25:85, SampleDate = .full_seq_dates(na.omit(serology$SampleDate)))
res = predict(model,newdata = new)
new = new %>% mutate(units = res)
p1 = ggplot(new, aes(x=SampleDate,y=age,fill=units, z=units))+geom_tile()+scale_fill_viridis_c(option = "magma")+metR::geom_contour2(colour="white")
p1 %>% ggrrr::gg_save_as(out("n-antibody-level"),size = std_size$half)
```

# N antibody positivity

* Not clear what level is of positivity
* Binomial model with logit link function (i.e. logistic regression with locally fitted polynomial on age and date)


```{r}
serology2 = serology %>% mutate(RocheN_pos = RocheN_outcome == "Positive") %>% filter(!is.na(RocheN_pos))
model = locfit::locfit(RocheN_pos ~ locfit::lp(age,SampleDate,nn=0.2,scale=TRUE),data = serology2, link="logit", family="qbinomial")

new = tidyr::crossing(age = 25:85, SampleDate = .full_seq_dates(na.omit(serology$SampleDate)))
res = predict(model,newdata = new)
new = new %>% mutate(positives = res)
p2 = ggplot(new, aes(x=SampleDate,y=age,fill=positives,z=positives))+geom_tile()+scale_fill_viridis_c(lim=c(0,1))+metR::geom_contour2(mapping = aes(z=positives,label = ..level..),colour="white", binwidth=0.1)
p2 %>% ggrrr::gg_save_as(out("n-antibody-positivity"),size = std_size$half)
```
# N antibody positivity by vaccination status (at time of sample)

* filters for any vaccine received prior to test.
* Not fully vaccinated.

```{r}
serology_unvac = serology2 %>% filter(is.na(firstvaccinationdate))
model = locfit::locfit(RocheN_pos ~ locfit::lp(age,SampleDate,nn=0.2,scale=TRUE),data = serology_unvac, link="logit", family="qbinomial")

new = tidyr::crossing(age = 25:85, SampleDate = .full_seq_dates(na.omit(serology$SampleDate)))
res = predict(model,newdata = new)
new = new %>% mutate(positives = res) %>% mutate(status = "Unvaccinated")

serology_vac = serology2 %>% filter(!is.na(firstvaccinationdate))
model2 = locfit::locfit(RocheN_pos ~ locfit::lp(age,SampleDate,nn=0.2,scale=TRUE),data = serology_vac, link="logit", family="qbinomial")

new2 = tidyr::crossing(age = 25:85, SampleDate = .full_seq_dates(na.omit(serology$SampleDate)))
res = predict(model2,newdata = new2)
new2 = new2 %>% mutate(positives = res) %>% mutate(status = "Vaccinated")

new3 = bind_rows(new,new2)

p3 = ggplot(new3, aes(x=SampleDate,y=age,fill=positives,z=positives))+geom_tile()+facet_wrap(vars(status))+
  scale_fill_viridis_c(lim=c(0,1))+metR::geom_contour2(mapping = aes(z=positives,label = ..level..),colour="white", binwidth=0.1)

p3 %>% ggrrr::gg_save_as(out("n-antibody-positivity-by-vaccination"),size = std_size$half)
```

# S antibody community level

```{r}
model2 = locfit::locfit(RocheS_units ~locfit::lp(age,SampleDate,nn=0.2,scale=TRUE),data = serology, link="log")
new2 = tidyr::crossing(age = 25:85, SampleDate = .full_seq_dates(na.omit(serology$SampleDate)))
res2 = predict(model2,newdata = new2)
new2 = new2 %>% mutate(units = res2)
p4 = ggplot(new2, aes(x=SampleDate,y=age,fill=units,z=units))+geom_tile()+scale_fill_viridis_c(option="magma")+metR::geom_contour2(colour="white")
p4 %>% ggrrr::gg_save_as(out("s-antibody-level"),size = std_size$half)
```
# S antibody positivity

* N.b. not sure what definition of positive is in this data.

```{r}
serology3 = serology %>% mutate(RocheS_pos = RocheS_outcome == "Positive") %>% filter(!is.na(RocheS_pos))
model = locfit::locfit(RocheS_pos ~ locfit::lp(age,SampleDate,nn=0.2,scale=TRUE),data = serology3, link="logit", family="qbinomial")

new = tidyr::crossing(age = 25:85, SampleDate = .full_seq_dates(na.omit(serology$SampleDate)))
res = predict(model,newdata = new)
new = new %>% mutate(positives = res)
p5 = ggplot(new, aes(x=SampleDate,y=age,fill=positives))+geom_tile()+
  scale_fill_viridis_c(lim=c(0,1))+metR::geom_contour2(mapping = aes(z=positives,label = ..level..),colour="white", binwidth=0.1)

p5 %>% ggrrr::gg_save_as(out("s-antibody-positivity"),size = std_size$half)
```

# Only Vaccine induced immunity

* S antibody positive and N antibody negative 
* Immunity that is solely conferred by vaccine

```{r}
serology4 = serology %>% mutate(RocheS_only = RocheS_outcome == "Positive" & RocheN_outcome == "Negative") %>% filter(!is.na(RocheS_only))
model = locfit::locfit(RocheS_only ~ locfit::lp(age,SampleDate,nn=0.2,scale=TRUE),data = serology4, link="logit", family="qbinomial")

new = tidyr::crossing(age = 25:85, SampleDate = .full_seq_dates(na.omit(serology$SampleDate)))
res = predict(model,newdata = new)
new = new %>% mutate(positives = res)
p6 = ggplot(new, aes(x=SampleDate,y=age,fill=positives,z=positives))+geom_tile()+
  scale_fill_viridis_c(lim=c(0,1))+metR::geom_contour2(mapping = aes(z=positives,label = ..level..),colour="white", binwidth=0.1)

p6 %>% ggrrr::gg_save_as(out("s-only-antibody-positivity"),size = std_size$half)
```

```{r}
# ggplot(serology %>% filter(!is.na(RocheS_units) & !is.na(RocheN_units)), aes(x=RocheS_units,y=RocheN_units, colour=secondvaccinationbrand))+geom_point()+scale_x_continuous(trans="log1p",breaks=ggrrr::breaks_log1p())+scale_y_continuous(trans="log1p",breaks=ggrrr::breaks_log1p())+
#   geom_vline(xintercept=c(275,2500),colour="red")
# 
# ggplot(serology %>% filter(!is.na(RocheS_units) & !is.na(RocheN_units)), aes(x=RocheS_units))+geom_density()+scale_x_continuous(trans="log1p")+
#   geom_vline(xintercept=c(275,2500),colour="red")
# ggplot(serology %>% filter(!is.na(RocheS_units) & !is.na(RocheN_units)), aes(x=RocheN_units))+geom_density()+scale_x_continuous(trans="log1p")
```


```{r}
# serology2 = serology %>% 
#   mutate(
#     dosesBeforeSample = case_when(
#       is.na(firstvaccinationdate) & is.na(secondvaccinationdate) & is.na(thirdvaccinationdate) ~ NA_character_,
#       SampleDate < firstvaccinationdate ~ "none",
#       is.na(secondvaccinationdate) | SampleDate < secondvaccinationdate ~ "one",
#       is.na(thirdvaccinationdate) | SampleDate < thirdvaccinationdate ~ "two",
#       TRUE ~ "three",
#     ) %>% factor(levels=c("none","one","two","three")),
#     lastDoseBeforeDate = case_when(
#       is.na(firstvaccinationdate) & is.na(secondvaccinationdate) & is.na(thirdvaccinationdate) ~ NA_Date_,
#       SampleDate < firstvaccinationdate ~ NA_Date_,
#       is.na(secondvaccinationdate) | SampleDate < secondvaccinationdate ~ firstvaccinationdate,
#       is.na(thirdvaccinationdate) | SampleDate < thirdvaccinationdate ~ secondvaccinationdate,
#       TRUE ~ thirdvaccinationdate
#     ),
#     ageCat = cut(age,breaks = c(35,65,75,Inf), labels = c("35-64","65-74","75+"),right = FALSE),
#     postDoseInterval = as.integer(SampleDate-lastDoseBeforeDate)
#   )
# 
# serology2 = serology2 %>% mutate(test = "RocheS_units", value=RocheS_units)
#   # pivot_longer(ends_with("units"), names_to="test")
```

```{r pressure, echo=FALSE}
# ggplot(serology2 %>% filter(!is.na(value) & !is.na(dosesBeforeSample) & !is.na(ageCat) & RocheN_units < 1), aes(x=postDoseInterval, y=value, colour=RocheN_units))+geom_point()+facet_wrap(vars(dosesBeforeSample,test),scales = "free_y")+scale_y_log10()
```


```{r}
# quantiles = c(0.025,0.25,0.5,0.75,0.975)
# tmp = serology2  %>% filter(!is.na(value) & !is.na(dosesBeforeSample) & !is.na(ageCat) & RocheN_units < 1 & !is.na(postDoseInterval) & !is.na(ageCat))
# 
# ggplot(tmp, aes(x=age,colour=secondvaccinationbrand))+geom_point(stat="count")+geom_line(stat="count")

```

```{r}
# any(is.na(tmp$postDoseInterval))
# 
# window=15
# 
# summaryLag = tmp %>% arrange(postDoseInterval) %>% 
#     group_by(dosesBeforeSample,ageCat) %>%
#     mutate(
#       rollingQuant = slider::slide_index(value, postDoseInterval, .before=window %/% 2,.after=window %/% 2,.complete = TRUE,.f = ~ enframe(quantile(.x, quantiles, na.rm=TRUE),value = "quant")),
#       rollingN = slider::slide_index_int(value, postDoseInterval, .before=window %/% 2,.after=window %/% 2,.complete = TRUE,.f = ~ length(.x)),
#       rollingMean = slider::slide_index_dbl(value, postDoseInterval, .before=window %/% 2,.after=window %/% 2,.complete = TRUE,.f = ~ mean(.x,na.rm=TRUE))
#     ) %>%
#     select(ageCat, dosesBeforeSample, postDoseInterval, rollingMean, rollingQuant, rollingN) %>%
#     unnest(rollingQuant) %>%
#     distinct()
# 
# ggplot(summaryLag%>% filter(rollingN>20), aes(x=postDoseInterval, y=rollingN))+geom_line()+facet_wrap(vars(dosesBeforeSample,ageCat),scales = "free_y")+scale_y_log10()
# 
# 
# peakImmunity = summaryLag %>% group_by(dosesBeforeSample,ageCat) %>% filter(rollingN>50 & name == "50%") %>% mutate(maxPostDoseInterval = max(postDoseInterval)) %>%    
#   # filter(log(quant)>0.95*max(log(quant))) %>%
#   filter(rollingMean>0.95*max(rollingMean)) %>% 
#   filter(postDoseInterval==min(postDoseInterval)) %>% select(ageCat, dosesBeforeSample, peakImmunityDays = postDoseInterval, medianPeak = quant, meanPeak = rollingMean, maxPostDoseInterval)
# 
# ggplot(summaryLag %>% filter(rollingN>30 & postDoseInterval<200), aes(x=postDoseInterval, y=quant, colour=name))+geom_line()+facet_grid(dosesBeforeSample~ageCat,scales = "free_y")+scale_y_log10()+
#   geom_vline(data = peakImmunity, aes(xintercept=peakImmunityDays), colour="grey")+
#   geom_point(data = peakImmunity, aes(x=peakImmunityDays, y = meanPeak), colour="black", size=1)
```


```{r}
# 
# serology3 = serology2 %>% inner_join(peakImmunity, by=c("dosesBeforeSample","ageCat")) %>% filter(postDoseInterval < maxPostDoseInterval) %>% mutate(prePeak = postDoseInterval <= peakImmunityDays, peakOffset = postDoseInterval-peakImmunityDays, normalisedValue = value/meanPeak)
# 
# tmp = serology3 %>% filter(dosesBeforeSample %in% c("two","three")) %>% mutate(
#   dosesBeforeSample = forcats::fct_drop(dosesBeforeSample),
#   ageCat = forcats::fct_drop(ageCat)
# )
# 
# 
# 
# tmp2 = tmp %>% filter(peakOffset >= 14 & peakOffset < 100 & (value < 2495 | value > 2505)) %>% group_by(dosesBeforeSample, ageCat, peakImmunityDays) %>% group_modify(function(d,g,...) {
#   d$init = 1
#   # fit <- glm(normalisedValue ~ peakOffset-1+offset(init), data=d, family=gaussian(link=log))
#   fit <- glm(value ~ peakOffset, data=d, family=gaussian(link=log))
#   predict = predict(fit,newdata=tibble(peakOffset=c(-14:150)),se.fit=TRUE)
#   return(tibble(
#     sample=list(d), fit= list(fit), 
#     predict=list(tibble(
#       peakOffset=c(-14:150),
#       value = predict$fit,
#       se = predict$se.fit
#     ))))
# })
# 
# tmp3 = tmp2 %>% mutate(coef = map(fit, ~ broom::tidy(.x)))
# tmp4 = tmp3 %>% select(dosesBeforeSample, coef) %>% unnest(coef)
# 
# 
# ggplot(tmp, aes(x=peakOffset, y=value))+geom_point()+geom_vline(colour="red",xintercept=0)+facet_grid(dosesBeforeSample~ageCat,scales = "free_y")+scale_y_log10()+
#   geom_ribbon(data=tmp2 %>% unnest(predict), aes(x=peakOffset, ymin =exp(value-1.96*se),ymax =exp(value+1.96*se)),fill="blue",alpha=0.2)+
#   geom_line(data=tmp2 %>% unnest(predict), aes(x=peakOffset, y =exp(value)),colour="blue",size=1)


```

```{r}
# ggplot(tmp4 %>% filter(term=="(Intercept)"), aes(x=dosesBeforeSample, colour=ageCat,y=exp(estimate), ymin=exp(estimate-1.96*std.error), ymax=exp(estimate+1.96*std.error)))+geom_point(position=position_dodge(width=0.6))+geom_errorbar(position=position_dodge(width=0.6),width=0.5)+
#   geom_hline(yintercept=1,colour="black")+ylab("peak units")+xlab("doses received")
# 
# ggplot(tmp4 %>% filter(term=="peakOffset"), aes(x=dosesBeforeSample, colour=ageCat,y=estimate, ymin=estimate-1.96*std.error, ymax=estimate+1.96*std.error))+geom_point(position=position_dodge(width=0.6))+geom_errorbar(position=position_dodge(width=0.6),width=0.5)+
#   geom_hline(yintercept=0,colour="black")+ylab("rate of change of log(units) per day")+xlab("doses received")
# 
# 
# tmp5 = tmp4 %>% filter(term=="peakOffset") %>% mutate(halflife = log(0.5)/estimate, halflife_lo = log(0.5)/(estimate+1.96*std.error), halflife_hi=log(0.5)/(estimate-1.96*std.error))
# 
# ggplot(tmp5, aes(x=dosesBeforeSample, colour=ageCat,y=halflife, ymin=halflife_lo, ymax=halflife_hi))+geom_point(position=position_dodge(width=0.6))+geom_errorbar(position=position_dodge(width=0.6),width=0.5)+
#   geom_hline(yintercept=0,colour="black")+ylab("half life (days)")+xlab("doses received")+coord_cartesian(ylim=c(-10,250))
# 
# 
# tmp4 %>% select(-statistic, -p.value) %>% pivot_wider(names_from = term,values_from=c(estimate, std.error)) %>% 
#   mutate(
#     halflife = log(0.5)/estimate_peakOffset, 
#     halflife_lo = log(0.5)/(estimate_peakOffset+1.96*std.error_peakOffset), 
#     halflife_hi=log(0.5)/(estimate_peakOffset-1.96*std.error_peakOffset),
#     `Half life (days)` = sprintf("%1.1f [%1.1f \u2015 %1.1f]", halflife, halflife_lo, halflife_hi)
#   ) %>% 
#   mutate(
#     `Peak level (Units)` = sprintf("%s [%s \u2015 %s]", 
#                                    signif(exp(`estimate_(Intercept)`),3), 
#                                    signif(exp(`estimate_(Intercept)`-1.96*`std.error_(Intercept)`),3), 
#                                    signif(exp(`estimate_(Intercept)`+1.96*`std.error_(Intercept)`),3))
#   ) %>%
#   select("Doses" = dosesBeforeSample, "Age category"=ageCat,"Days to peak"=peakImmunityDays,`Peak level (Units)`, `Half life (days)`) %>%
#   ggrrr::hux_tidy(rowGroupVars = vars(Doses,`Age category`), colGroupVars = vars()) %>%
#   clipr::write_clip()


```